An Automatic Verification Framework for Software Configurations

### 2016-05-05 OSDI/intro.tex
config: new fig and motivation

While many efforts have been proposed 
to check, troubleshoot, diagnose and repair configuration errors%
~\cite{attariyan10automating,
su07autobash, whitaker04configuration}, 
offering automatic verification to configuration files -- like what we
did for verifying regular programs -- is still highly
desirable~\cite{wang04automatic, zhang14encore, xu15systems}.
Nevertheless, the main obstacles to the configuration verification are:
1) entries in configuration files are untyped entries, 2) there
is no explicit structure policy for the entries in configuration
files, and 3) there are surprisingly few rules specifying the
entries' constraints.

In order to overcome these obstacles,
researchers propose statistical analysis and learning based approaches%
~\cite{wang04automatic, zhang14encore, yuan11context}. 
These efforts build checking policies by learning a sample 
data set, rather than explicitly specifying entries' types or rules.
In particular, for each entry in a certain configuration file, 
if it deviates from the common values used in a large collection
of configurations (\ie, the training data set), it is typically
suspected as a potential configuration error.
However, because these learning 
efforts {\em either} are limited to simplistic 
configuration errors (\eg, type errors and syntax errors), 
{\em or} heavily rely on template-based inference~\cite{zhang14encore}, 
many sophisticated configuration errors, 
\eg, hard to be templated in practice, cannot be detected.
For example, if {\tt extension = mysql.so} appears 
before {\tt extension = recode.so} in PHP configuration file,  
it would lead to a crash error, since the correct ordering 
should be {\tt extension = recode.so} before 
{\tt extension = mysql.so}~\cite{yin11anempirical};
however, such an error cannot be detected by existing
learning efforts, since it is hard to build a template for this
error~\cite{xu15systems}.

@@

Our \app prototype still has a few limitations,
\eg, cannot handle errors occurred in execution time.
Nevertheless, we believe \app may suggest a practical path
toward automatic and language-based configuration verification.
To summarize, this tool paper makes the following contributions:

>>>>

While many efforts have been proposed 
to check, troubleshoot, diagnose and repair configuration errors%
~\cite{attariyan10automating,
su07autobash, whitaker04configuration}, 
offering automatic verification to configuration files -- like what we
did for verifying regular programs -- is still highly
desirable~\cite{wang04automatic, zhang14encore, xu15systems}.
Nevertheless, the main obstacles to the automatic 
configuration verification are:
1) entries in configuration files are untyped entries, 2) there
is no explicit structure policy for the entries in configuration
files, and 3) there are surprisingly few rules specifying the
entries' constraints.

In order to overcome these obstacles,
researchers have proposed 
statistical analysis and learning based approaches%
~\cite{wang04automatic, zhang14encore, yuan11context}. 
These efforts build checking policies by learning a sample 
data set, rather than explicitly specifying entries' types or rules.
In particular, for each entry in a certain configuration file, 
if it deviates from the common values used in a large collection
of configurations (\ie, the data set for training), it is typically
suspected as a potential configuration error.
However, because these learning 
efforts {\em either} are limited to simplistic 
configuration errors (\eg, type errors and syntax errors), 
{\em or} heavily rely on template-based inference~\cite{zhang14encore}, 
many sophisticated configuration errors, 
\eg, hard to be templated in practice, cannot be detected.
For example, if {\tt extension = mysql.so} appears 
before {\tt extension = recode.so} in PHP configuration file,  
it would lead to a crash error, since the correct ordering 
should be {\tt extension = recode.so} before 
{\tt extension = mysql.so}~\cite{yin11anempirical};
however, such an error cannot be detected by existing
learning efforts, since it is hard to build a template for this
error~\cite{xu15systems}.

@@

Our \app prototype still has a few limitations:
for example, we cannot handle configuration errors that can be only
triggered in system execution time.
Nevertheless, we believe \app may suggest a practical path
toward automatic and modular language-based configuration verification.
To summarize, this tool paper makes the following contributions:

### 2016-05-05 OSDI/motiv.tex
config: new fig and motivation

\begin{lstlisting}[language=C]
extension = mysql.so
...
extension = recode.so
\end{lstlisting}

@@

\begin{lstlisting}[language=C]
ORDERING ERROR: Expected "extension" "recode.so"
BEFORE "extension" "mysql.so"
\end{lstlisting}

@@

\begin{lstlisting}[language=C]
key_buffer_size = 384M
max_heap_table_size = 128M
max_connections = 64
thread_cache_size = 8
...
sort_buffer_size = 32M
join_buffer_size = 32M
read_buffer_size = 32M
read_rnd_buffer_size = 8M
...
\end{lstlisting}

This user complained his MySQL's load is very high and the website's
respond speed is very slow.
In this case, {\tt key\_buffer\_size} is used by all the threads
cooperatively while {\tt join\_buffer} and {\tt sort\_buffer} are 
created by each thread for private use; thus, the maximum amount
of used key buffer, \ie, {\tt key\_buffer\_size} should be larger than 
{\tt join|sort\_buffer\_size} * {\tt max\_connections}. 
Clearly, in the above example, it is not, so that this misconfiguration
causes MySQL loads very slow.

\app is able to check such fine-grained value correlation problem.
If we run \app on this configuration file, \app  would return:

\begin{lstlisting}[language=C]
INTEGER RELATION ERROR:
Expected "key_buffer_size">="max_connections"*"sort_buffer_size"
\end{lstlisting}

Note that existing effort, EnCore, aims to detect value correlation 
errors. However, different from \app, EnCore can only detect simple
correlation, \eg, {\tt mysql.max\_persistent} < {\tt max\_connections},
rather than the fine-grained computation-based value correlation
done by \app.

\para{Example~3: Entry missing errors.} 
If a user wants to use OpenLDAP to enable her directory access
protocol, she needs to use the password policy overlay. This is usually
done through the following entries in the OpenLDAP configuration file:

\begin{lstlisting}[language=C]
include schema/ppolicy.schema
overlay ppolicy
\end{lstlisting}

When using the password policy overlay in OpenLDAP, 
we have to first include the related schema.
Leaving out the ``include'' statement will cause the failure of 
this LDAP. Running \app on such a misconfiguration file would return:

\begin{lstlisting}[language=C]
MISSING KEYWORD ERROR: Expected "overlay" "ppolicy"
in the same file: "include" "schema/ppolicy.schema"
\end{lstlisting}

>>>>

\begin{lstlisting}[language=C, xleftmargin=.01\textwidth]
extension = mysql.so
...
extension = recode.so
\end{lstlisting}

@@

\begin{lstlisting}[language=C, xleftmargin=.01\textwidth]
ORDERING ERROR: Expected "extension" "recode.so"
BEFORE "extension" "mysql.so"
\end{lstlisting}

@@

\begin{lstlisting}[language=C, xleftmargin=.01\textwidth]
key_buffer_size = 384M
max_heap_table_size = 128M
max_connections = 64
thread_cache_size = 8
...
sort_buffer_size = 32M
join_buffer_size = 32M
read_buffer_size = 32M
read_rnd_buffer_size = 8M
...
\end{lstlisting}

This user complained his MySQL's load is very high and the website's
respond speed is very slow.
In this case, {\tt key\_buffer\_size} is used by all the threads
cooperatively while {\tt join\_buffer} and {\tt sort\_buffer} are 
created by each thread for private use; thus, the maximum amount
of used key buffer, \ie, {\tt key\_buffer\_size} should be larger than 
{\tt join|sort\_buffer\_size} * {\tt max\_connections}. 
Clearly, in the above example, it does not hold, 
so that this misconfiguration causes MySQL loads very slow.

We call the above situation as a {\em fine-grained value correlation
error}, which is a more challenging case than 
{\em value correlation error} reported
by some existing investigations~\cite{yin11anempirical, zhang14encore}.
A typical value correlation error
means one entry's value should have a certain correlation with
another entry's value. For example, in MySQL,
the value of {\tt max\_connections} should be higher than
{\tt mysql.max\_persistent}. Although some work, \eg,
EnCore~\cite{zhang14encore}, can detect the normal value
correlation case, 
none of existing efforts can detect the fine-grained value correlation
errors, which ``hides'' an equation-aware correlation, as shown in
the above example.
As mentioned by Xu {\em et al.}~\cite{xu15hey},
detecting the fine-grained value correlation errors 
present a much more challenging
task than normal value correlation problems.

To the best of our knowledge, \app is the first effort that
is able to check such fine-grained value correlation problem.
If we run \app on this configuration file, \app  would return:

\begin{lstlisting}[language=C, xleftmargin=.01\textwidth]
INTEGER RELATION ERROR:
Expected "key_buffer_size" >= "max_connections" * "sort_buffer_size"
\end{lstlisting}

Note that existing effort, EnCore, aims to detect value correlation 
errors. However, different from \app, EnCore can only detect simple
correlation, \eg, {\tt mysql.max\_persistent} < {\tt max\_connections},
rather than the fine-grained computation-based value correlation,
like {\tt key\_buffer\_size >= max\_connections * sort\_buffer\_size}
identified by \app.

\para{Example~3: Entry missing errors.} 
Many critical system outages result from the fact that important
entry was missing in the configuration file. 
We call such a problem as {\em entry missing error}.
Below is a real-world entry missing error example,
when a user wants to use OpenLDAP to enable her directory access
protocol, she needs to use the password policy overlay. This is usually
done through the following entries in the OpenLDAP configuration file:

\begin{lstlisting}[language=C, xleftmargin=.01\textwidth]
include schema/ppolicy.schema
overlay ppolicy
\end{lstlisting}

When using the password policy overlay in OpenLDAP, 
users have to first include the related schema.
Leaving out the ``include'' entry, 
which have been left by many users~\cite{yin11anempirical}, 
will cause the failure of this LDAP. 
If this user has \app, she can run \app on such a misconfiguration file,
and \app would return:

\begin{lstlisting}[language=C, xleftmargin=.01\textwidth]
MISSING KEYWORD ERROR: Expected "overlay" "ppolicy"
in the same file: "include" "schema/ppolicy.schema"
\end{lstlisting}
